\documentclass{article}
\input{Core Files/preamble}

\title{Solid State Physics Notes}
\author{Alex Hardie}

\begin{document}

\maketitle
\tableofcontents
\newpage


\chaptertitle{Chapter 1: What is Condensed Matter Physics?}{Updated Jan 21\textsuperscript{st},2025}
\input{Chapter Assets/Ch. 1 Assets/SimonSolidStateCh1Notes}


\chaptertitle{Chapter 2: Boltzmann, Einstein, and Debye}{Updated Jan 21\textsuperscript{st}, 2025}
\input{Chapter Assets/Ch. 2 Assets/SimonSolidStateCh2Notes}

\chaptertitle{Chapter 3: Drude Theory}{Updated Jan 21\textsuperscript{st}, 2025}
\input{Chapter Assets/Ch. 3 Assets/SimonSolidStateCh3Notes}

\chaptertitle{Chapter 4: Sommerfeld (Free Electron) Model}{}
\input{Chapter Assets/Ch. 4 Assets/SimonSolidStateCh4Notes}

\chaptertitle{Chapter 5: The Periodic Table}{}
\input{Chapter Assets/Ch. 5 Assets/SimonSolidStateCh5Notes}

\chaptertitle{Chapter 6: Chemical Bonding}{}
\input{Chapter Assets/Ch. 6 Assets/SimonSolidStateCh6Notes}

\chaptertitle{Chapter 7: Types of Matter}{}
\input{Chapter Assets/Ch. 7 Assets/SimonSolidStateCh7Notes}

\chaptertitle{Chapter 8: One-Dimensional Model of Compressibility, Sound, and Thermal Expansion}{}
\input{Chapter Assets/Ch. 8 Assets/SimonSolidStateCh8Notes}

\chaptertitle{Chapter 9: Vibrations of a One-Dimensional Monoatomic Chain}{}
\input{Chapter Assets/Ch. 9 Assets/SimonSolidStateCh9Notes}

\chaptertitle{Chapter 10: Vibrations of a One-Dimensional Monoatomic Chain}{}
\input{Chapter Assets/Ch. 10 Assets/SimonSolidStateCh10Notes}

\chaptertitle{Chapter 11: Tight Binding Chain (Interlude and Preview)}{}
\input{Chapter Assets/Ch. 11 Assets/SimonSolidStateCh11Notes}

\chaptertitle{Chapter 12: Crystal Structure}{}
\input{Chapter Assets/Ch. 12 Assets/SimonSolidStateCh12Notes}

\chaptertitle{Chapter 13: Reciprocal LAttice, Brillouin Zone, Waves in Crystals}{}
\input{Chapter Assets/Ch. 13 Assets/SimonSolidStateCh13Notes}

\chaptertitle{Chapter 14: Wave Scattering by Crystals}{}
\input{Chapter Assets/Ch. 14 Assets/SimonSolidStateCh14Notes}

\chaptertitle{Chapter 15: Electrons in a Periodic Potential}{}
\input{Chapter Assets/Ch. 15 Assets/SimonSolidStateCh15Notes}

\chaptertitle{Chapter 16: Insulator, Semiconductor, or Metal}{}
\input{Chapter Assets/Ch. 16 Assets/SimonSolidStateCh16Notes}

\chaptertitle{Chapter 17: Semiconductor Physics}{}
\input{Chapter Assets/Ch. 17 Assets/SimonSolidStateCh17Notes}

\chaptertitle{Chapter 18: Semiconductor Devices}{}
\input{Chapter Assets/Ch. 18 Assets/SimonSolidStateCh18Notes}

\chaptertitle{Chapter 19: Semiconductor Devices}{}
\input{Chapter Assets/Ch. 19 Assets/SimonSolidStateCh19Notes}

\chaptertitle{Chapter 20: Ferro-, Antiferr-, and Ferri-Magnetism}{}
\input{Chapter Assets/Ch. 20 Assets/SimonSolidStateCh20Notes}

\chaptertitle{Chapter 21: Domains and Hysteresis}{}
\input{Chapter Assets/Ch. 21 Assets/SimonSolidStateCh21Notes}

\chaptertitle{Chapter 22: Domains and Hysteresis}{}
\input{Chapter Assets/Ch. 22 Assets/SimonSolidStateCh22Notes}

\chaptertitle{Chapter 23: Domains and Hysteresis}{}
\input{Chapter Assets/Ch. 23 Assets/SimonSolidStateCh23Notes}

\end{document}